---
title: "NBA player clustering"
output: html_document
---
#Load all necessary packages
tidyverse for wrangling and manipulating data
nbastatR and ballr to webscrape nba statistics
ggfortify and GGally for autoplot data visualizations
```{r packages, message = F, warning = F}
library(tidyverse)
library(nbastatR)
library(ballr)
library(ggfortify)
library(GGally)
```

Obtain raw data and wrangle so it can be analyzed
```{r obtain and wrangle data, message = F, warning = F}
teamplayerstats <- teams_players_stats(seasons = 2019, types = "player")

playerbox <- teams_players_stats(seasons = 2017:2019,
                               types = "player",
                               tables = c("general", "hustle", "shot locations"),
                               measures = c("Base", "Advanced"),
                               modes = "PerGame",
                               defenses = "Overall",
                               assign_to_environment = F)

seasonTables <- list()
for (tablenum in 1:nrow(playerbox)){
  seasonTables[[tablenum]] <- playerbox$dataTable[[tablenum]] %>% mutate(season = playerbox$slugSeason[tablenum])
}
playerbox$dataTable = seasonTables

playerboxstats <- left_join(
  
  bind_rows(lapply(playerbox$dataTable[1:3], function(df){df %>% select(10, 76, 14, 20:22, 34:35, 41:43, 45:53)})),
  
  bind_rows(lapply(playerbox$dataTable[4:6], function(df){df %>% select(4, 31, 9:11, 14:15, 17, 20)})),
  
  by = c("namePlayer" = "namePlayer", "season" = "season")
  
) %>%
  
  left_join(
    
    bind_rows(lapply(playerbox$dataTable[7:9], function(df){df %>% select(11, 37, 16:17, 19:20, 22:23, 25:26, 28:29, 31:32, 34:35)})),
    
    by = c("namePlayer" = "namePlayer", "season" = "season")
    
  ) %>% 
  
  left_join(
    
    bind_rows(lapply(playerbox$dataTable[10:12], function(df){df %>% select(10, 87, 16:18, 20:24, 50, 53, 56, 58)})),
    
    by = c("namePlayer" = "namePlayer", "season" = "season")
  )

playerboxstats <- playerboxstats %>%
  select(-23) %>% 
  filter(
    namePlayer %in% (playerboxstats %>% group_by(namePlayer) %>% summarise(totalgp = sum(gp)) %>% filter(totalgp >= 50))$namePlayer
  )

pergamecols <- c(4, 7, 9, 10:19, 21:24, 26:27, 29, 31, 33, 35, 37, 39)
othercols <- setdiff(4:52, pergamecols)

pergame <- playerboxstats %>% select(pergamecols)
pergame <- pergame %>% 
  mutate_all(function(col){col * playerboxstats$gp}) %>% 
  cbind(playerboxstats[,1:3]) %>% 
  group_by(namePlayer) %>% 
  summarise_if(is.numeric, sum) 

pergame1 <- pergame %>% select(-c(1, 27)) %>% 
  mutate_all(function(col){col / pergame$gp}) %>% 
  cbind(pergame %>% select(1))

other <- playerboxstats %>% select(1, othercols) %>% 
  group_by(namePlayer) %>% 
  summarise_if(is.numeric, mean)

playerboxstats <- left_join(other, pergame1, by = "namePlayer") %>% 
  mutate(nameAbrev = paste(substring(str_split_fixed(namePlayer, " ", n = 2)[,1], 1, 1), ". ", str_split_fixed(namePlayer, " ", n = 2)[,2], sep = "")) %>% 
  select(c(51, 2:50)) %>% 
  na.omit()
```


```{r}
pboxmat <- as.matrix(playerboxstats[,c(44:49, 29, 30, 35, 19)])
rownames(pboxmat) <- playerboxstats$nameAbrev
pboxout <- prcomp(pboxmat, scale = T, center = T)
autoplot(pboxout, label = T, shape = F, loadings = T, loadings.label = T)
pboxout$rotation
help(autoplot)
```


```{r Player possessions, message = F, warning = F}
playersynergy <- synergy(seasons = 2013:2019, result_types = "player", set_types = c("offensive"), categories = c("Transition", "Isolation", "PRBallHandler", "PRRollman", "Postup",  "Spotup", "Handoff", "Cut", "OffScreen", "OffRebound", "Misc"))

synplayer <- lapply(playersynergy$dataSynergy, function(df){
  dfnew <- df %>% 
    mutate(season = playersynergy$slugSeason[sapply(playersynergy$dataSynergy, all.equal, current = df) == "TRUE"]) %>% 
    select(season, namePlayer, pctFrequency)
  colnames(dfnew)[3] <- paste(playersynergy$categorySynergy[sapply(playersynergy$dataSynergy, all.equal, current = df) == "TRUE"], "pctFrequency")
  dfnew
}) %>% 
  
  reduce(full_join, by = c("season" = "season", "namePlayer" = "namePlayer")) %>% 
  
  mutate(
    `Transition pctFrequency` = as.numeric(substring(str_replace_all(paste(`Transition pctFrequency.x`, `Transition pctFrequency.y`, `Transition pctFrequency`, sep = ""), "[NA]", ""), 1, 8)),
    
    `Isolation pctFrequency` = as.numeric(substring(str_replace_all(paste(`Isolation pctFrequency.x`, `Isolation pctFrequency.y`, `Isolation pctFrequency.x.x`, `Isolation pctFrequency.y.y`, sep = ""), "[NA]", ""), 1, 8)),
    
    `PRBallHandler pctFrequency` = as.numeric(substring(str_replace_all(paste(`PRBallHandler pctFrequency.x`, `PRBallHandler pctFrequency.y`, `PRBallHandler pctFrequency.x.x`, `PRBallHandler pctFrequency.y.y`, sep = ""), "[NA]", ""), 1, 8)),
    
    `PRRollman pctFrequency` = as.numeric(substring(str_replace_all(paste(`PRRollman pctFrequency.x`, `PRRollman pctFrequency.y`, `PRRollman pctFrequency.x.x`, `PRRollman pctFrequency.y.y`, sep = ""), "[NA]", ""), 1, 8)),
    
    `Postup pctFrequency` = as.numeric(substring(str_replace_all(paste(`Postup pctFrequency.x`, `Postup pctFrequency.y`, `Postup pctFrequency.x.x`, `Postup pctFrequency.y.y`, sep = ""), "[NA]", ""), 1, 8)),
    
    `Spotup pctFrequency` = as.numeric(substring(str_replace_all(paste(`Spotup pctFrequency.x`, `Spotup pctFrequency.y`, `Spotup pctFrequency.x.x`, `Spotup pctFrequency.y.y`, sep = ""), "[NA]", ""), 1, 8)),
    
    `Handoff pctFrequency` = as.numeric(substring(str_replace_all(paste(`Handoff pctFrequency.x`, `Handoff pctFrequency.y`, `Handoff pctFrequency.x.x`, `Handoff pctFrequency.y.y`, sep = ""), "[NA]", ""), 1, 8)),
    
    `Cut pctFrequency` = as.numeric(substring(str_replace_all(paste(`Cut pctFrequency.x`, `Cut pctFrequency.y`, `Cut pctFrequency.x.x`, `Cut pctFrequency.y.y`, sep = ""), "[NA]", ""), 1, 8)),
    
    `OffScreen pctFrequency` = as.numeric(substring(str_replace_all(paste(`OffScreen pctFrequency.x`, `OffScreen pctFrequency.y`, `OffScreen pctFrequency.x.x`, `OffScreen pctFrequency.y.y`, sep = ""), "[NA]", ""), 1, 8)),
    
    `OffRebound pctFrequency` = as.numeric(substring(str_replace_all(paste(`OffRebound pctFrequency.x`, `OffRebound pctFrequency.y`, `OffRebound pctFrequency.x.x`, `OffRebound pctFrequency.y.y`, sep = ""), "[NA]", ""), 1, 8)),
    
    `Misc pctFrequency` = as.numeric(substring(str_replace_all(paste(`Misc pctFrequency.x`, `Misc pctFrequency.y`, `Misc pctFrequency.x.x`, `Misc pctFrequency.y.y`, sep = ""), "[NA]", ""), 1, 8))
         ) %>% 
  #left_join(teamdeets$dataTable[[1]] %>% select(nameTeam, slugTeam), by = "nameTeam") %>% 
  
  #mutate(seasonTeam = paste(substring(season, 3, 7), slugTeam)) %>% 
    
  #filter(season %in% c("2016-17", "2017-18", "2018-19")) %>%
  
  #left_join(passing, by = c("season" = "Year", "nameTeam" = "Team")) %>% 
  
  select(namePlayer, `Transition pctFrequency`, `Isolation pctFrequency`, `PRBallHandler pctFrequency`, `PRRollman pctFrequency`, `Postup pctFrequency`, `Spotup pctFrequency`, `Handoff pctFrequency`, `Cut pctFrequency`, `OffScreen pctFrequency`, `OffRebound pctFrequency`, `Misc pctFrequency`)

synplayer[is.na(synplayer)] <- 0

synplayer <- synplayer %>% 
  group_by(namePlayer) %>% 
  
  summarise(`Transition pctFrequency` = mean(`Transition pctFrequency`),
            `Isolation pctFrequency` = mean(`Isolation pctFrequency`),
            `PRBallHandler pctFrequency` = mean(`PRBallHandler pctFrequency`),
            `PRRollman pctFrequency` = mean(`PRRollman pctFrequency`),
            `Postup pctFrequency` = mean(`Postup pctFrequency`),
            `Spotup pctFrequency` = mean(`Spotup pctFrequency`),
            `Handoff pctFrequency` = mean(`Handoff pctFrequency`),
            `Cut pctFrequency` = mean(`Cut pctFrequency`),
            `OffScreen pctFrequency` = mean(`OffScreen pctFrequency`),
            `OffRebound pctFrequency` = mean(`OffRebound pctFrequency`),
            `Misc pctFrequency` = mean(`Misc pctFrequency`)
            ) %>% 
  filter(namePlayer %in% playershots$namePlayer)
```
  
```{r Dataframe Dictionary, message = F, warning = F}
teamdeets <- teams_details(all_teams = T)
teamabrev <- teamdeets[[2]][[1]] %>% select(nameTeam, slugTeam)
teamplayerstats <- teams_players_stats(seasons = 2019, types = "team")
df_dict_nba_teams <- df_nba_team_dict 
```

```{r NBA team data: Plays, message = F, warning = F}
teamplays <- synergy(seasons = 2016:2019,
                   result_types = c("team"),
                   season_types = c("Regular Season"),
                   set_types = c("offensive"),
                   categories = c("Transition", "Isolation", "PRBallHandler", "PRRollman", "Postup", "Spotup", "Handoff", "Cut", "OffScreen", "OffRebound", "Misc"),
                   results = 500,
                   assign_to_environment = TRUE,
                   return_wide = FALSE,
                   return_message = TRUE)
```

```{r NBA team data: Shots, message = F, warning = F}
teamshots17 <- teams_shots(all_active_teams = T, seasons = 2017)
teamshots18 <- teams_shots(all_active_teams = T, seasons = 2018)
teamshots19 <- teams_shots(all_active_teams = T, seasons = 2019)

teamshots <- rbind(teamshots17, teamshots18, teamshots19) %>% 
  mutate(nameTeam = case_when(nameTeam == "LA Clippers" ~ "Los Angeles Clippers",
                              TRUE ~ nameTeam)) %>%
  group_by(slugSeason, nameTeam, typeAction) %>% 
  summarise(n = n()) %>% 
  mutate(freq = n / sum(n),
         
         typeAction = str_to_lower(typeAction),
         
         broadAction = case_when(typeAction %in% c("alley oop dunk shot", 
                                                   "alley oop layup shot", 
                                                   "running alley oop dunk shot", 
                                                   "running alley oop layup shot") ~ "Alley oop freq",
                                 
                                 typeAction %in% c("cutting dunk shot",
                                                   "cutting finger roll layup shot",
                                                   "cutting layup shot") ~ "Cut freq",
                                 
                                 str_detect(typeAction, "driving") | typeAction %in% c("finger roll layup shot",
                                                                                       "floating jump shot",
                                                                                       "layup shot",
                                                                                       "reverse layup shot",
                                                                                       "running finger roll layup shot",
                                                                                       "running hook shot",
                                                                                       "running layup shot",
                                                                                       "running reverse layup shot",
                                                                                       "running dunk shot",
                                                                                       "running reverse dunk shot") ~ "Drive freq",
                                 typeAction %in% c("dunk shot",
                                                   "reverse dunk shot") ~ "Plain dunk freq",
                                 
                                 typeAction %in% c("fadeaway bank shot",
                                                   "fadeaway jump shot",
                                                   "step back bank jump shot",
                                                   "step back jump shot",
                                                   "turnaround fadeaway bank jump shot",
                                                   "turnaround fadeaway shot",
                                                   "turnaround jump shot") ~ "Separation jump shot freq",
                                 
                                 typeAction %in% c("hook bank shot",
                                                   "hook shot",
                                                   "turnaround bank hook shot",
                                                   "turnaround bank shot",
                                                   "turnaround hook shot") ~ "Post up freq",
                                 
                                 typeAction %in% c("jump bank shot",
                                                   "jump shot") ~ "Plain jump shot freq",
                                 
                                 typeAction %in% c("pullup bank shot",
                                                   "pullup jump shot",
                                                   "running pull-up jump shot",
                                                   "running jump shot") ~ "Pullup/running jump shot freq",
                                 
                                 typeAction %in% c("putback dunk shot",
                                                   "putback layup shot",
                                                   "tip dunk shot",
                                                   "tip layup shot") ~ "Putback/Tip freq"
                                 )
         ) %>% 
  group_by(slugSeason, nameTeam, broadAction) %>% 
  summarise(freq = sum(freq)) %>% 
  spread(broadAction, freq) %>% 
  select(-12) %>%
  
  left_join(teamabrev, by = "nameTeam") %>% 
  mutate(seasonTeam = paste(substring(slugSeason, 3, 7), slugTeam)) %>% 
  ungroup() %>% 
  select(-slugTeam, -slugSeason, -nameTeam) %>% 
  select(10, 2:9)
```

```{r NBA team data: Passing}
teampassing <- read.csv("C:/Users/bchan/Documents/NBA/20117-2019 Team Passing.csv") %>% 
  select(1,2,7,13) %>% 
  mutate(Team = case_when(as.character(Team) == "LA Clippers" ~ "Los Angeles Clippers",
                          TRUE ~ as.character(Team))) %>% 
  left_join(teamabrev, by = c("Team" = "nameTeam")) %>% 
  mutate(seasonTeam = paste(substring(Year, 3, 7), slugTeam))
```
Only keeping passes made and adjusted assists because the other stats are measuring the same things

```{r NBA team data: Box and Advanced Stats, message = F, warning = F}
teambox <- teams_players_stats(seasons = 2017:2019,
                               types = "team",
                               tables = c("general", "hustle", "shot locations"),
                               measures = c("Base", "Advanced"),
                               modes = "PerGame",
                               defenses = "Overall",
                               assign_to_environment = F)

seasonTables <- list()
for (tablenum in 1:nrow(teambox)){
  seasonTables[[tablenum]] <- teambox$dataTable[[tablenum]] %>% mutate(season = teambox$slugSeason[tablenum])
}
teambox$dataTable = seasonTables

teamboxstats <- left_join(
  
  bind_rows(lapply(teambox$dataTable[1:3], function(df){df %>% select(10, 67, 32, 33, 18:20, 34, 38:40, 43, 44, 46, 47:50)})),
  
  bind_rows(lapply(teambox$dataTable[4:6], function(df){df %>% select(4, 24, 7, 8, 10:12, 16)})),
  
  by = c("nameTeam" = "nameTeam", "season" = "season")
  
) %>%
  
  left_join(
    
    bind_rows(lapply(teambox$dataTable[7:9], function(df){df %>% select(11, 34, 14, 15, 17, 18, 20, 21, 23, 24, 26, 27, 29, 30, 32, 33)})),
    
    by = c("nameTeam" = "nameTeam", "season" = "season")
    
  ) %>% 
  
  left_join(
    
    bind_rows(lapply(teambox$dataTable[10:12], function(df){df %>% select(10, 56, 14:16, 18:20, 35, 37, 39, 40, 43)})),
    
    by = c("nameTeam" = "nameTeam", "season" = "season")
    
  ) %>% 
  mutate(nameTeam = case_when(as.character(nameTeam) == "LA Clippers" ~ "Los Angeles Clippers",
                          TRUE ~ as.character(nameTeam))) %>% 
  left_join(teamabrev, by = c("nameTeam" = "nameTeam")) %>% 
  mutate(seasonTeam = paste(substring(season, 3, 7), slugTeam)) %>% 
  select(-nameTeam, -season) %>% 
  select(49, 1:47)
```

Tables to join
teamshots
teamboxstats

```{r NBA team data: Compile}
NBAteams <- lapply(teamplays$dataSynergy, function(df){
  dfnew <- df %>% 
    mutate(season = teamplays$slugSeason[sapply(teamplays$dataSynergy, all.equal, current = df) == "TRUE"]) %>% 
    select(season, nameTeam, pctFrequency)
  colnames(dfnew)[3] <- paste(teamplays$categorySynergy[sapply(teamplays$dataSynergy, all.equal, current = df) == "TRUE"], "pctFrequency")
  dfnew
}) %>% 
  
  reduce(full_join, by = c("season" = "season", "nameTeam" = "nameTeam")) %>% 
  
  mutate(
    `Transition pctFrequency` = as.numeric(substring(str_replace_all(paste(`Transition pctFrequency.x`, `Transition pctFrequency.y`, `Transition pctFrequency.x.x`, `Transition pctFrequency.y.y`, sep = ""), "[NA]", ""), 1, 8)),
    
    `Isolation pctFrequency` = as.numeric(substring(str_replace_all(paste(`Isolation pctFrequency.x`, `Isolation pctFrequency.y`, `Isolation pctFrequency.x.x`, `Isolation pctFrequency.y.y`, sep = ""), "[NA]", ""), 1, 8)),
    
    `PRBallHandler pctFrequency` = as.numeric(substring(str_replace_all(paste(`PRBallHandler pctFrequency.x`, `PRBallHandler pctFrequency.y`, `PRBallHandler pctFrequency`, sep = ""), "[NA]", ""), 1, 8)),
    
    `PRRollman pctFrequency` = as.numeric(substring(str_replace_all(paste(`PRRollman pctFrequency.x`, `PRRollman pctFrequency.y`, `PRRollman pctFrequency.x.x`, `PRRollman pctFrequency.y.y`, sep = ""), "[NA]", ""), 1, 8)),
    
    `Postup pctFrequency` = as.numeric(substring(str_replace_all(paste(`Postup pctFrequency.x`, `Postup pctFrequency.y`, `Postup pctFrequency.x.x`, `Postup pctFrequency.y.y`, sep = ""), "[NA]", ""), 1, 8)),
    
    `Spotup pctFrequency` = as.numeric(substring(str_replace_all(paste(`Spotup pctFrequency.x`, `Spotup pctFrequency.y`, `Spotup pctFrequency.x.x`, `Spotup pctFrequency.y.y`, sep = ""), "[NA]", ""), 1, 8)),
    
    `Handoff pctFrequency` = as.numeric(substring(str_replace_all(paste(`Handoff pctFrequency.x`, `Handoff pctFrequency.y`, `Handoff pctFrequency.x.x`, `Handoff pctFrequency.y.y`, sep = ""), "[NA]", ""), 1, 8)),
    
    `Cut pctFrequency` = as.numeric(substring(str_replace_all(paste(`Cut pctFrequency.x`, `Cut pctFrequency.y`, `Cut pctFrequency.x.x`, `Cut pctFrequency.y.y`, sep = ""), "[NA]", ""), 1, 8)),
    
    `OffScreen pctFrequency` = as.numeric(substring(str_replace_all(paste(`OffScreen pctFrequency.x`, `OffScreen pctFrequency.y`, `OffScreen pctFrequency.x.x`, `OffScreen pctFrequency.y.y`, sep = ""), "[NA]", ""), 1, 8)),
    
    `OffRebound pctFrequency` = as.numeric(substring(str_replace_all(paste(`OffRebound pctFrequency.x`, `OffRebound pctFrequency.y`, `OffRebound pctFrequency`, sep = ""), "[NA]", ""), 1, 8)),
    
    `Misc pctFrequency` = as.numeric(substring(str_replace_all(paste(`Misc pctFrequency.x`, `Misc pctFrequency.y`, `Misc pctFrequency`, sep = ""), "[NA]", ""), 1, 8))
         ) %>% 
  left_join(teamdeets$dataTable[[1]] %>% select(nameTeam, slugTeam), by = "nameTeam") %>% 
    
  filter(season %in% c("2016-17", "2017-18", "2018-19")) %>%
  
  left_join(teampassing, by = c("season" = "Year", "nameTeam" = "Team")) %>% 
  
  select(season, nameTeam, seasonTeam, `Transition pctFrequency`, `Isolation pctFrequency`, `PRBallHandler pctFrequency`, `PRRollman pctFrequency`, `Postup pctFrequency`, `Spotup pctFrequency`, `Handoff pctFrequency`, `Cut pctFrequency`, `OffScreen pctFrequency`, `OffRebound pctFrequency`, `Misc pctFrequency`, PASSES.MADE, ADJUSTED.ASSISTS) %>% 
  
  left_join(teamshots, by = "seasonTeam") %>% 
  
  left_join(teamboxstats, by = "seasonTeam")

colnames(NBAteams)
```

not much here
```{r Team PCA}
teammat <- as.matrix(NBAteams[,4:71])
rownames(teammat) <- NBAteams$seasonTeam
teamout <- prcomp(teammat, scale = T, center = T)
autoplot(teamout, label = T, shape = F)

pr.var <- teamout$sdev^2
pve = pr.var / sum(pr.var)
par(mfrow = c(1,2))
plot(pve, xlab = "Principal Component", ylab = "Proportion of Variance Explained", ylim = c(0,1), type = 'b')

plot(cumsum(pve), xlab = "Principal Component", ylab = "Proportion of Variance Explained", ylim = c(0,1), type = 'b')

library(GGally)
km <- kmeans(teamout$x[,1:60], centers = 3)
View(as.data.frame(km$cluster) %>% mutate(seasonTeam = rownames(as.data.frame(km$cluster))) %>% arrange(-`km$cluster`))

huh1 <- as.data.frame(teamout$rotation[,1:3]) %>% 
  select(1)
huh1 <- huh1 %>% 
  mutate(thang = rownames(huh1)) 
huh1 <- huh1 %>% 
  arrange(-PC1)

huh2 <- as.data.frame(teamout$rotation[,1:3]) %>% 
  select(2)
huh2 <- huh2 %>% 
  mutate(thang = rownames(huh2)) 
huh2 <- huh2 %>% 
  arrange(-PC2)

huh3 <- as.data.frame(teamout$rotation[,1:3]) %>% 
  select(3)
huh3 <- huh3 %>% 
  mutate(thang = rownames(huh3)) 
huh3 <- huh3 %>% 
  arrange(-PC3)
```
13


