---
title: "spodifye"
output: html_document
---

```{r Reading in JSON data, message = F, warning = F}
library(rjson)
library(tidyverse)
library(spotifyr)
library(ggfortify)
library(factoextra)
library(GGally)

streamJson <- "C:/Users/bchan/Documents/my_spotify_data/MyData/StreamingHistory.json"
streamList <- fromJSON(paste(readLines(streamJson), collapse = ""))
streamdf <- bind_rows(lapply(streamList, as.data.frame))
```

```{r favtracks}
favtracks <- streamdf %>% 
  group_by(trackName, artistName) %>% 
  summarise(totalTime = sum(msPlayed)) %>% 
  arrange(-totalTime)
```


```{r}
Sys.setenv(SPOTIFY_CLIENT_ID = 'b933e8011ebe4d68a6dcb80fa40faa36')
Sys.setenv(SPOTIFY_CLIENT_SECRET = '0f3629522e8249f5bca055bd2ab7c7a5')

token <- get_spotify_access_token()
```

```{r, message = F, warning = F}
trackartist <- streamdf %>% 
  distinct(trackName, artistName) %>% 
  mutate(trartist = paste(trackName, artistName, sep = "|||"))


search_id <- function(trartist){
  trackName = unlist(strsplit(trartist, "|||", 2))[1]
  artistName = unlist(strsplit(trartist, "|||", 2))[2]
  search <- search_spotify(q = paste("track:", trackName, " artist:", artistName, sep = ""), type = "track", limit = 1, authorization = token)
  if(nrow(search) == 0){return("still nah")}
  return(search$id[1])
}

ids1 <- c()
for (tracknum in 1:length(trackartist$trartist)){
  ids1[tracknum] = search_id(trackartist$trartist[tracknum])
}

trackartist$ids1 <- ids1
trackartist <- trackartist %>% 
  mutate(id = case_when(trackName == "Fried Rice (with JID)" ~ search_id("Fried Rice (with JID)|||Bas"),
                          TRUE ~ ids1)
  ) %>% 
  select(-ids1, -trartist) %>% 
  filter(id != "still nah")

#write.csv(trackartist, file = "Tracks with IDs.csv", row.names = F)
```


```{r}
featrowlist <- list()
low <- 1
high <- 100
for (i in 1:((nrow(trackartist) %/% 100) + 1)){
  featrowlist[[i]] <- low:high
  low <- high + 1
  high <- low + 99
  if (high > nrow(trackartist)){
    high <- nrow(trackartist)
  }
}

features <- bind_rows(lapply(featrowlist, function(rows){
  get_track_audio_features(trackartist$id[rows], token)
})) %>% 
  select(c(13, 1:11, 17:18))
```

```{r}
trackfeatures <- left_join(trackartist, features, by = "id") %>% 
  left_join(favtracks, by = c("trackName" = "trackName", "artistName" = "artistName")) %>% 
  filter(totalTime >= duration_ms) %>% 
  arrange(-totalTime)

mostlistened <- trackfeatures %>% 
  mutate(trackName = case_when(str_detect(trackName, "\\(") ~ str_sub(trackName, start = 1, end = str_locate(trackName, "\\(")[,1] - 2),
                               str_detect(trackName, "\\|")  ~ str_sub(trackName, start = 1, end = str_locate(trackName, "\\|")[,1] - 2),
                               TRUE ~ trackName)
  )

mostartists <- mostlistened %>% 
  group_by(artistName) %>% 
  summarise_if(is.numeric, mean)

colnames(mostartists)[2:15] <- c("Danceability", "Energy", "Key", "Loudness", "Mode", "Speechiness", "Acousticness", "Instrumentalness", "Liveness", "Valence", "Tempo", "Song Length", "Time Signature", "Total Time Listened")

mostlistened %>% arrange(-liveness)
  
#mostlistened %>% arrange(-key)
```

FINAL CLUSTERS
Cluster artists
  key to duration
  PC1 - high loudness, low everything else
  PC2 - high everything else except accousticness, low instrumentalness
  PC3 - high accousticness, low everything else except key
  PC4 - high mode, low liveness
  PC5 - high speechiness and lowkey accousticness, low everything else
kmeans for PC1 to PC5 with 4 clusters/centers
Clusters
  Rap 
    high speechiness (highest by far), liveness, valence, loudness
    low instrumentalness
  General Turn up (singing)
    high loudness, valence, liveness
    low instrumentalness and speechiness
  Instrumentals
    high instrumentalness (highest by far), accousticness
    low loudness, speechiness, and valence
  Sad Boi
    high accousticness (highest by far)
    low energy, loudness, speechiness, liveness, valence, and tempo

Mode indicates the modality of a track, the type of scale from which its melodic content is derived. Major is 1 and minor is 0.

```{r}
mat <- as.matrix((mostartists %>% na.omit())[,c(2:13)])
rownames(mat) <- (mostartists %>% na.omit())$artistName
pcout <- prcomp(mat, scale = T, center = F)

biplot <- autoplot(pcout, 
         loadings = T, 
         loadings.label = T,
         loadings.label.colour = "#1DB954",
         loadings.colour = "#1DB954",
         shape = F, 
         label = T,
         label.size = 3,
         loadings.label.vjust = 1.4,
         loadings.label.size = 4) + xlab("Principal Component 1") + ylab("Principal Component 2")
biplot
ggsave(filename = "Spotify Biplot.png", plot = biplot, width = 9, device = NULL, path = "C:/Users/bchan/Documents/Final Presentation")
```

```{r}
pcvariance <- function(pcoutt, pcnum){
  variances <- (pcoutt$sdev) ^ 2
  paste(signif(((variances[pcnum] / sum(variances)) * 100), 5), "%", sep = "")
}

pcvariance(pcout, 1)
```


```{r}
loadings <- data.frame(
  
  loading = c(pcout$rotation[,1], pcout$rotation[,2], pcout$rotation[,3], pcout$rotation[,4], pcout$rotation[,5], pcout$rotation[,6]),
  
  PCnum = c(rep.int(1, 12), rep.int(2, 12), rep.int(3, 12), rep.int(4, 12), rep.int(5, 12), rep.int(6, 12)),
  
  playtype = rep(colnames(mostartists)[2:13], 6)
) %>% 
  mutate(sign = case_when(loading < 0 ~ "n",
                          loading > 0 ~ "p"),
         absloading = abs(loading))

loadingsplot <- function(PCnumber){
  ggplot((loadings %>% filter(PCnum == PCnumber)), aes(x = reorder(playtype, loading), y = loading, fill = absloading)) + 
  geom_col() + 
  theme(text = element_text(size = 20), axis.text.x = element_text(angle = 50, hjust = 1), legend.position = "none") + 
  scale_fill_gradient(low = "#0d4a22", high = "#28de67") + 
  xlab("Audio Feature") + ylab("Loading Value")
}

loadings1 <- loadingsplot(1)
loadings2 <- loadingsplot(2)
loadings3 <- loadingsplot(3)
loadings4 <- loadingsplot(4)
loadings5 <- loadingsplot(5)
loadingsplot(6)

ggsave(filename = "Spotify PC1 Loadings.png", plot = loadings1, device = NULL, width = 13, path = "C:/Users/bchan/Documents/Final Presentation")
ggsave(filename = "Spotify PC2 Loadings.png", plot = loadings2, device = NULL, width = 13, path = "C:/Users/bchan/Documents/Final Presentation")
ggsave(filename = "Spotify PC3 Loadings.png", plot = loadings3, device = NULL, width = 13, path = "C:/Users/bchan/Documents/Final Presentation")
ggsave(filename = "Spotify PC4 Loadings.png", plot = loadings4, device = NULL, width = 13, path = "C:/Users/bchan/Documents/Final Presentation")
ggsave(filename = "Spotify PC5 Loadings.png", plot = loadings5, device = NULL, width = 13, path = "C:/Users/bchan/Documents/Final Presentation")
```

```{r}
km <- kmeans(pcout$x[,1:5], centers = 4)
halp <- as.data.frame(km$cluster) %>% mutate(team = row.names(as.data.frame(km$cluster))) %>% arrange(km$cluster) %>% select(team, `km$cluster`)
ggplot(, aes(x = pcout$x[,1], y = pcout$x[,2])) + geom_point(aes(col = as.factor(km$cluster))) + xlab("First PC") + ylab("Second PC") + scale_color_discrete(name = "Cluster") + geom_point(aes(x = km$centers[,1], y = km$centers[,2], pch = factor(1:4)))
mostartists$cluster <- km$cluster
```

```{r}
trackfeatures %>% filter(artistName == "Monte Booker")
View(mostartists %>% filter(as.numeric(cluster) == 4))
mostartistscluster <- mostartists %>% mutate(cluster = as.factor(cluster))
clusters <- autoplot(pcout, 
                     data = mostartistscluster,
                     colour = "cluster",
                     label = T, 
                     shape = F, 
                     loadings = T, 
                     loadings.label = T,
                     loadings.label.colour = "black",
                     loadings.label.hjust = 0.5,
                     loadings.label.vjust = 1.4,
                     loadings.colour = "black",
                     label.size = 3,
                     loadings.label.size = 4) + 
  xlab("Principal Component 1") +
  ylab("Principal Component 2") +
  scale_color_discrete(name = "Cluster", labels = c("Turn Up", "Rap", "Instrumental", "Turn Down"))
clusters
ggsave(filename = "Spotify Clustering w Loadings.png", plot = clusters, device = NULL, width = 11, path = "C:/Users/bchan/Documents/Final Presentation")

#C77CFF purple
#F8766D orange
#00BFC4 blue
#7CAE00 green

clustersNoLoad <- autoplot(pcout, 
                     data = mostartistscluster,
                     colour = "cluster",
                     label = T, 
                     shape = F, 
                     loadings = F, 
                     loadings.label = F,
                     loadings.label.colour = "black",
                     loadings.label.hjust = 0.5,
                     loadings.label.vjust = 1.4,
                     loadings.colour = "black",
                     label.size = 3,
                     loadings.label.size = 4) + 
  xlab("Principal Component 1") +
  ylab("Principal Component 2") +
  scale_color_discrete(name = "Cluster", labels = c("Turn Up", "Rap", "Instrumental", "Turn Down"))
ggsave(filename = "Spotify Clustering wo Loadings.png", plot = clustersNoLoad, device = NULL, width = 11, path = "C:/Users/bchan/Documents/Final Presentation")
```

```{r}
View(mostartistscluster %>% filter(as.numeric(cluster) == 3))
```


```{r}
mostartistscluster <- mostartistscluster %>% 
  mutate(cluster = as.numeric(cluster)) %>% 
  select(-`Time Signature`)

avgcols <- colnames(mostartistscluster)[2:13]

one <- (mostartistscluster %>% filter(cluster == 1)) %>% 
  mutate(numart = n()) %>% 
  select(-artistName) %>% 
  summarise_all(list(~mean(.), ~sum(.))) %>% 
  select(c(1:12, 14, 28))

two <- (mostartistscluster %>% filter(cluster == 2)) %>% 
  mutate(numart = n()) %>% 
  select(-artistName) %>% 
  summarise_all(list(~mean(.), ~sum(.))) %>% 
  select(c(1:12, 14, 28))

three <- (mostartistscluster %>% filter(cluster == 3)) %>% 
  mutate(numart = n()) %>% 
  select(-artistName) %>% 
  summarise_all(list(~mean(.), ~sum(.))) %>% 
  select(c(1:12, 14, 28))

avgfeatures <- (mostartistscluster %>% filter(cluster == 4)) %>% 
  mutate(numart = n()) %>% 
  select(-artistName) %>% 
  summarise_all(list(~mean(.), ~sum(.))) %>% 
  select(c(1:12, 14, 28)) %>% 
  bind_rows(three, two , one)

colnames(avgfeatures) <- c("Danceability", "Energy", "Key", "Loudness", "Mode", "Speechiness", "Acousticnesss", "Instrumentalness", "Liveness", "Valence", "Tempo", "Song Length", "Cluster", "Total Time Listened")

nonscale <- avgfeatures %>% select(Cluster, `Total Time Listened`)

avgfeatures <- avgfeatures %>% 
  select(-Cluster, -`Total Time Listened`) %>% 
  scale() %>% 
  cbind(nonscale) %>% 
  gather(feature, score, -Cluster)
```

```{r}
featureorder <- c("Loudness", "Energy", "Valence", "Tempo", "Danceability", "Speechiness", "Liveness", "Song Length", "Key", "Mode", "Acousticnesss", "Instrumentalness")
featureplot <- function(clusternum, color){
  ggplot((avgfeatures %>% filter(Cluster == clusternum & !str_detect(feature, "Total"))), aes(x = feature, y = score)) + 
    geom_col(fill = color) + 
    theme(text = element_text(size = 20), axis.text.x = element_text(angle = 50, hjust = 1)) +
   scale_x_discrete(limits = featureorder) +
    ylim(-1.5, 1.5) + 
    xlab("Audio Feature") +
    ylab("Standardized Average")
}

features1 <- featureplot(1, "#F8766D")
features2 <- featureplot(2, "#7CAE00")
features3 <- featureplot(3, "#00BFC4")
features4 <- featureplot(4, "#C77CFF")

ggsave(filename = "Spotify Turn Up Features.png", plot = features1, device = NULL, width = 13, path = "C:/Users/bchan/Documents/Final Presentation")
ggsave(filename = "Spotify Rap Features.png", plot = features2, device = NULL, width = 13, path = "C:/Users/bchan/Documents/Final Presentation")
ggsave(filename = "Spotify Instrumental Features.png", plot = features3, device = NULL, width = 13, path = "C:/Users/bchan/Documents/Final Presentation")
ggsave(filename = "Spotify Turn Down Features.png", plot = features4, device = NULL, width = 13, path = "C:/Users/bchan/Documents/Final Presentation")
```

```{r}
write.csv(mostartistscluster, file = "C:/Users/bchan/Documents/Final Presentation/spotify artsits clustered.csv", row.names = F)
```

